---
title: "session3"
author: "Claire Madden"
date: "1/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# set global options here
```

### 1. Attach necessary packages

```{r, include=FALSE}
library(tidyverse)
library(sf)
library(tmap)
library(here)
library(leaflet)

```

### 2. Read in California eco-regions data
```{r}

ca_eco <- read_sf(here("session_3_materials"), layer = "ca_eco") %>% 
  select(US_L3NAME) %>% 
  rename(region = US_L3NAME) %>% 
  st_simplify(dTolerance = 100) %>% 
  st_transform(crs = 4326)

# rmapshaper:: ms_simplify(), geos::gSimplify will both do the same thing as st_simplify but also account for the fact that all polygons should be connected


#to check coordinate system - st_crs()
  #set coordinate system - st_transfrom()
# if projection doesn't exist at all you can use st_crs()

plot(ca_eco)

```


### 3. Read in CA counties polygons

```{r}

ca_counties <- read_sf(here("session_3_materials"), layer = "california_county_shape_file") 

st_crs(ca_counties) = 4326


```


### 4. Read in CA dams locations data

```{r}

ca_dams <- read_sf(here("session_3_materials"), layer = "California_Jurisdictional_Dams")


```


### 5. Make a map of CA with all three datasets!

```{r}

ggplot(data = ca_counties) +
  geom_sf(color = "black", 
          size = 0.1) +
  geom_sf(data = ca_eco, 
          aes(fill = region),
          alpha = 0.5, 
          color = "NA", 
          show.legend = FALSE) +
  geom_sf(data = ca_dams,
          size = 0.5,
          alpha = 0.5) +
  theme_minimal()

```


#### 6. Just check out dams in the Sierra Nevada eco-region

Join together the dam data with the eco-region data, filter to only include Sierra Nevada eco-region, retain only the attribute for Dam_Name, and make a map!

```{r}

# First: join intersected dams in Sierra Nevada eco-region
sn <- ca_eco %>% 
  filter(region == "Sierra Nevada") %>% 
  st_join(ca_dams)

ggplot(sn) +
  geom_sf(data = ca_counties, fill = "wheat3", color = "NA") +
  geom_sf(fill = "lemonchiffon4", color = "NA") +
  geom_point(aes(x = Longitude, y = Latitude), size = 0.5, color = "red4") +
  theme_void() +
  coord_sf(datum=NA) +
  labs(x = "", y = "", title = "CA Dams in Sierra Nevada Eco-Region")

```



#### 7. Just explore SB County eco-regions (clipping)

Can plot just pieces using st_intersection (for example, if we only want to plot eco-regions in Santa Barbara County), and crop graphing space with coord_sf() limits.

```{r}

# Get just SB county
sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara")

# Clip eco-region spatial data to intersection with SB county:
eco_clip <- st_intersection(ca_eco, sb)

```

Plot it!
```{r}
ggplot(eco_clip) +
  geom_sf(data = ca_counties, fill = "gray90", color = "gray80", size = 0.2) + # First add gray California
  geom_sf(aes(fill = region), color = "NA") + # ...then add eco-regions (clipped)
  scale_fill_manual(values = c("darkolivegreen2","darkolivegreen","gold2")) + # Change color scheme
  coord_sf(xlim = c(-121,-119), ylim = c(33.5,35.5)) + # Crop plotting area
  geom_point(aes(x = -119.6982, y = 34.4208), size = 2) + # Add a point for SB City
  geom_text(x = -119.6982, y = 34.35, label = "Santa Barbara") + # Then add a label for it
  theme_minimal() + # Update theme
  theme(legend.position = c(0.5,0.15)) +# Move the legend
  labs(x = "", y = "", title = "Santa Barbara County Eco-Regions")
```

Hooray! Clipped polygons.


### 7. Create an interactive map with tmap

```{r}

sb_clip_tmap <- tm_basemap("Esri.WorldImagery")+
  tm_shape(eco_clip)+
  tm_fill("region", palette = c("orange", "purple", "yellow"),
          alpha = 0.5)
  
tmap_mode("view")
sb_clip_tmap
```



### 8. How to create an sf object from lat lon recordings (ex: stored in an excel file)
make a moc dataset using tribble()

```{r}

my_data <- tribble(
  ~id, ~lon, ~lat,
  "tiger", -119.4, 34.35,
  "lion", -119.41, 34.39,
  "bear", -119.43, 34.38
)

# convert this to an sf object :
# have to indicate columns where lon and lat are stored, in that order

animals_sf <- st_as_sf(my_data, coords = c("lon", "lat"),
                       crs = 4326)



```



Make a map using tmap

```{r}

animal_map <- tm_basemap("Stamen.Watercolor") +
  tm_shape(animals_sf)+ #this layer just tells r what data to use
  tm_dots(labels = "id", col = "grey", size = 0.5)


animal_map

```

### 9. Chloropleth of dam counts

```{r}

intersection <- st_intersection(x = ca_dams, y = ca_counties)

dams_per_county <- intersection %>% 
  group_by(NAME) %>% 
  tally()



# merge count data with county geometry to be able to map counties


ca_tot <- ca_counties %>% 
  st_join(dams_per_county) %>% 
  select(NAME.x, n) %>% 
  rename(name = NAME.x)


ca_tot_zero <- ca_tot %>%
  replace_na(list(n = 0))



ggplot()+
  geom_sf(data = ca_tot_zero,
          aes(fill = n),
          size = 0.2,
          color = "white")+
  scale_fill_continuous(low = "yellow", high = "red")

```







